---
title: "Raster Pre-Processing"
author: "Oliver Atwood"
output: html_document
date: "`r Sys.Date()`"
---
# Overview
<br>
This r markdown script pre-processes raster data used in the Hotspot Stoplight Project for visualization and cluster analysis. The datasets used in this project are as follows:

* LAND COVER CHANGE PROBABILITY - [Clark Labs](https://env1.arcgis.com/arcgis/rest/services/Land_Cover_Vulnerability_2050/ImageServer0)
* BIODIVERSITY INTACTNESS INDEX - [Impact Observatory](https://gee-community-catalog.org/projects/bii/)
* CLIMATE RISK INDEX - Hotspot Stoplight Team: [Nissim Lebovits](http://nlebovits.github.io/)
* GRIDDED 2020 POPULATION - [Global Human Settlements Layer](https://ghsl.jrc.ec.europa.eu/download.php?ds=bu)
* URBAN LAND COVER PROBABILITY - Hotspot Stoplight Team: [Oliver Atwood](https://www.linkedin.com/in/oliveratwood), using classified data from [Tristan Grupp](https://scholar.google.com/citations?user=yX3Ed8EAAAAJ&hl=en)
<br>
<br>

* Each of these raster datasets are loaded into R and their Coordinate Systems, Resolution, and Extent are standardized. NA values are then removed and each dataset is standardized.
<br>
* Four of these standardized datasets are then combined to produce two 'Interaction Rasters', using the following formulas:
<br>
+ Bio_Loss_Risk = LCC_Probability * BII
<br>
*Bio_Loss_Risk emphasizes areas of high land cover change probability and high biodiversity intactness.*
<br>
+ Climate_Impact_Risk = Climate_Hazards * population
<br>
*Climate_Impact_Risk emphasizes areas of high climate hazard probability and high human population.*
<br>
+ Urban Land Cover Probability is a dataset generated through a random forest model trained on a range of physiographic factors to predict the likelihood of a given cell of a land cover raster to be urban in 2033.
<br>

* Bio_Loss_Risk and Climate_Impact_Risk were both normalized and Bio_Loss_Risk, Climate_Impact_Risk, and Urban Land Cover Probability were exported for clustering analysis.
<br>
* These three datasets were then added together and the resultant raster dataset was normalized to produce a 'Stoplight for Urban Development'.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE)
rm(list=ls())
```

```{r, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide', echo=FALSE}
library(sf)
library(raster)
library(dplyr)
library(readr)
library(terra)
library(spdep)
library(rasterVis)
library(lattice)
library(gridExtra)
library(viridis)
library(classInt)
library(sp)
library(RColorBrewer)
library(grDevices)
library(stars)
```

## Setting Parameters
```{r, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide', echo=FALSE}
coordinate_system <- 4326
# Target resolution in m
resolution <- 30
# resolution <- 60
# resolution <- 100
# Conversion to degrees
resolution <- (resolution/30)/3600

```

## Loading Datasets
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE, results = 'hide'}
# Base directory
inputs <- '/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop5c_30m/'

outputs <- "/Users/oliveratwood/Documents/GitHub/HotspotStoplight/VisualProduction/OUTPUTS/Crop5c_30m/tiffs"

plot_dir <- '/Users/oliveratwood/Documents/GitHub/HotspotStoplight/VisualProduction/OUTPUTS/Crop5c_30m/plots'

# Load the raster datasets
lc_2023 <- raster(paste0(inputs, 'lc_2023.tif'))
lc_2050 <- raster(paste0(inputs, 'lc_2050.tif'))
UrbanProbability <- raster(paste0(inputs, 'Built.tif'))
OtherProbability <- raster(paste0(inputs, 'Other.tif'))

# # List of file names
# filenames <- c('lc_2023.tif', 'lc_2050.tif', 'Built.tif', 'Other.tif')
# var_names <- c('lc_2023', 'lc_2050', 'UrbanProbability', 'OtherProbability')
# 
# # Loop through each file name and assign its full path to a dynamically named variable
# for (i in 1:length(filenames)) {
#   # Generate the full path for the current file
#   full_path <- file.path(inputs, filenames[i])
# 
#   # Dynamically create a variable with the name from var_names and assign the full path to it
#   assign(var_names[i], full_path)
# }


BII <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Biodiversity/data/outputs/BiodiversityIntactness_2020.tif")
Flood_Probability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/outputs/data-costa_rica-outputs-flood_prob.tif")
Heat_Probability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/outputs/data-costa_rica-outputs-predicted_median_top5_costa_rica.tif")
population <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/vulnerability/data/inputs/High_Resolution_Settlement_Layer_crop2.tif")
GDP <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/vulnerability/data/inputs/GDP_2018_USD_Millions_Crop2.tif")


# Read the Boundary files
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5a.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5b.geojson")
cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5c.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop6.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop4.geojson")

# # PRESENT LAND COVER
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/lc_2023.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# lc_2023 <- raster(temp)
# 
# # PREDICTED LAND COVER
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/lc_2050.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# lc_2050 <- raster(temp)
# 
# # LAND COVER CHANGE PROBABILITY
# # CLARK LABS
# # url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/Land_Cover_Change_Probs_Clark.tif"
# 
# # URBAN
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/Built.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# UrbanProbability <- raster(temp)
# 
# # OTHER
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/Other.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# OtherProbability <- raster(temp)
# 
# # LCC_Probability <- OtherProbability + UrbanProbability
# # LCC_Probability <- OtherProbability
# 
# # BIODIVERSITY INTACTNESS INDEX
# url <- "https://github.com/HotspotStoplight/Biodiversity/raw/main/data/outputs/BiodiversityIntactness_2020.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# BII <- raster(temp)
# 
# # CLIMATE HAZARDS
# # Flood Hazard Probability
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/outputs/data-San_Jose-outputs-flood_prob.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# Flood_Probability <- raster(temp)
# 
# # Heat Hazard Probability
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/outputs/data-costa_rica-outputs-predicted_hot_days_costa_rica.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# Heat_Probability <- raster(temp)
# 
# # POPULATION
# # url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/Pop_2020_Crop4.tif"
# # temp <- tempfile(fileext = ".tif")
# # download.file(url, temp, method = "auto")
# # population <- raster(temp)
# 
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/vulnerability/data/inputs/High_Resolution_Settlement_Layer_crop2.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# population <- raster(temp)
# 
# # GDP
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/vulnerability/data/inputs/GDP_2018_USD_Millions_Crop2.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# GDP <- raster(temp)

# List of raster variables
raster_list = list(BII, Flood_Probability, Heat_Probability, population, UrbanProbability, OtherProbability, GDP)
lc_rasters = list(lc_2023, lc_2050)

```
## Define Normalization Function (MIN-MAX)
```{r}
norm <- function(raster) {
  # Calculate the minimum and maximum values of the raster
  min_val <- minValue(raster)
  max_val <- maxValue(raster)
  
  # Perform min-max normalization
  # Note: Use overlay or calc function for raster operations
  normalized_raster <- calc(raster, function(x) {
    # Apply the min-max normalization formula
    (x - min_val) / (max_val - min_val)
  })
  
  # Handle any potential division by zero or other anomalies
  normalized_raster[is.na(normalized_raster) | is.infinite(normalized_raster)] <- NA
  
  return(normalized_raster)
}
```

## Unify Raster Coordinate Systems, Resolutions, Extents, and remove NA Values
```{r}
unify_noncategorical <- function(raster_list, cropbox, resolution) {
  # Extract extent, CRS, and calculate rows and columns from cropbox
  bbox <- st_bbox(cropbox)
  target_extent <- as(extent(bbox), "Extent")
  target_crs <- st_crs(cropbox)$proj4string
  nrows <- ceiling((bbox["ymax"] - bbox["ymin"]) / resolution)
  ncols <- ceiling((bbox["xmax"] - bbox["xmin"]) / resolution)
  
  # Create an empty template raster with the specified extent, resolution, and CRS
  template_raster <- raster(nrows=nrows, ncols=ncols, xmn=bbox["xmin"], xmx=bbox["xmax"], ymn=bbox["ymin"], ymx=bbox["ymax"], crs=target_crs)
  standardized_rasters <- list() # Initialize list for standardized rasters
  
  for (i in seq_along(raster_list)) {
    current_raster <- raster_list[[i]]
    # Reproject to match the target CRS
    if (!compareCRS(crs(current_raster), target_crs)) {
      current_raster <- projectRaster(current_raster, crs = target_crs)
    }
    # Resample to match target resolution and extent, using bilinear interpolation
    current_raster <- resample(current_raster, template_raster, method = "bilinear")
    # Crop to match the target extent
    current_raster <- crop(current_raster, target_extent)
    # Replace NA values with 0
    current_raster[is.na(getValues(current_raster))] <- 0
    # Add to list of standardized rasters
    standardized_rasters[[i]] <- current_raster
  }
  
  names(standardized_rasters) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "GDP")
  
  return(standardized_rasters)
}

raster_list <- unify_noncategorical(raster_list, cropbox, resolution)

```

```{r}
unify_categorical <- function(raster_list, cropbox, resolution) {
  # Extract extent, CRS, and calculate rows and columns from cropbox
  bbox <- st_bbox(cropbox)
  target_extent <- as(extent(bbox), "Extent")
  target_crs <- st_crs(cropbox)$proj4string
  nrows <- ceiling((bbox["ymax"] - bbox["ymin"]) / resolution)
  ncols <- ceiling((bbox["xmax"] - bbox["xmin"]) / resolution)
  
  # Create an empty template raster with the specified extent, resolution, and CRS
  template_raster <- raster(nrows=nrows, ncols=ncols, xmn=bbox["xmin"], xmx=bbox["xmax"], ymn=bbox["ymin"], ymx=bbox["ymax"], crs=target_crs)
  standardized_rasters <- list() # Initialize list for standardized rasters
  
  for (i in seq_along(raster_list)) {
    current_raster <- raster_list[[i]]
    # Reproject to match the target CRS
    if (!compareCRS(crs(current_raster), target_crs)) {
      current_raster <- projectRaster(current_raster, crs = target_crs)
    }
    # Resample to match target resolution and extent, using nearest neighbor interpolation
    current_raster <- resample(current_raster, template_raster, method = "ngb")
    # Crop to match the target extent
    current_raster <- crop(current_raster, target_extent)
    # Replace NA values with 0
    current_raster[is.na(getValues(current_raster))] <- 0
    # Add to list of standardized rasters
    standardized_rasters[[i]] <- current_raster
  }
  
  names(standardized_rasters) <- c("lc_2023", "lc_2050")
  
  return(standardized_rasters)
}

lc_rasters <- unify_categorical(lc_rasters, cropbox, resolution)

```

## Filter Land Cover Type Probability by Predicted Change
```{r}
# pull lc_2023 out of raster list for easier access
lc_2023 <- lc_rasters$lc_2023

# Initialize LCC_Probability with lc_2023 dimensions and values
LCC_Probability <- lc_2023
values(LCC_Probability) <- 0  # Set all initial values to 0

# Assigning probabilities by predicted land cover change class
# Forest to Other
LCC_Probability[values(lc_2023) == 1] <- values(OtherProbability)[values(lc_2023) == 1]

# Other to Urban
LCC_Probability[values(lc_2023) == 2] <- values(UrbanProbability)[values(lc_2023) == 2]

# Add LCC_Probability to the raster_list
raster_list$LCC_Probability <- LCC_Probability

# Optionally, if you want to keep the names consistent and in a specific order, you can rename all at once
names(raster_list) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "GDP", "LCC_Probability")

```

```{r, fig.width=10}
# Create a new raster for UrbanExProbability based on the conditions
UrbanExProbability <- raster(UrbanProbability)

# Set UrbanExProbability to 0 where lc_2023 is 3
UrbanExProbability[values(lc_2023) == 3] <- 0

# Set UrbanExProbability to the UrbanProbability value where lc_2023 is not 3
UrbanExProbability[values(lc_2023) != 3] <- values(UrbanProbability)[values(lc_2023) != 3]

# Now plot UrbanExProbability
plot(UrbanExProbability, col=inferno(100))

# # Add LCC_Probability to the raster_list
raster_list$UrbanExProbability <- UrbanExProbability
# 
# # Optionally, if you want to keep the names consistent and in a specific order, you can rename all at once
names(raster_list) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "GDP", "LCC_Probability", "UrbanExProbability")
```


```{r, fig.width=8}
plot(LCC_Probability, col=inferno(100))
```
## Normalize all rasters
```{r}
# Apply the normalization function to each raster layer in the list
normalized_layers <- lapply(raster_list, function(layer) {
  norm(layer)
})
raster_list <- normalized_layers
names(raster_list) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "GDP", "LCC_Probability", "UrbanExProbability")
```

## Plot Input Rasters
```{r, fig.width=10}
# Convert the Raster list to a stack
raster_stack <- stack(normalized_layers)
plot(raster_stack, col=viridis::viridis(100))
```
## Generate Interaction Rasters
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
#URBAN PROBABILITY
UrbanProbability <- raster_list$UrbanProbability

#BIO RISK
# Multiply LCC_Probability and BII
Bio_Loss_Risk <- norm(raster_list$BII * raster_list$LCC_Probability)
# Bio_Loss_Risk <- norm(raster_list$BII * raster_list$OtherProbability)

#CLIMATE RISK
#Generate Climate Risk Raster
Climate_Hazards <- norm(raster_list$Flood_Probability * raster_list$Heat_Probability)

#Generate vulnerability-adjusted population raster
GDP_Inverse <- (1 - raster_list$GDP)
Population_Vulnerability <- norm(raster_list$population * GDP_Inverse)

# Multiply Vulnerability-adjusted population and Climate_Hazards
Climate_Impact_Risk <- norm(raster_list$population * Climate_Hazards)
Climate_Impact_Risk_v2 <- norm(Population_Vulnerability * Climate_Hazards) # needs a higher-resolution socioeconomic dataset


```

## Overlay Rasters to Produce 'Stoplight For Urban Development'
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
# Weighted Overlay Calculation
# Hotspot_Stoplight <- norm(UrbanProbability - Bio_Loss_Risk - Climate_Impact_Risk)
Hotspot_Stoplight <- norm(UrbanProbability - Bio_Loss_Risk - Climate_Impact_Risk - raster_list$BII)
# Hotspot_Stoplight_Expansion <- norm(raster_list$UrbanExProbability - Bio_Loss_Risk - Climate_Impact_Risk)
```

```{r}
library(raster)

# Assuming raster_list$UrbanExProbability, Bio_Loss_Risk, and Climate_Impact_Risk are RasterLayer objects

# Perform the initial calculation
temp_calculation <- raster_list$UrbanExProbability - Bio_Loss_Risk - Climate_Impact_Risk

# Create a function that will be used to normalize the raster values
norm_calculation <- function(x) {
  min_val <- min(x, na.rm = TRUE)
  max_val <- max(x, na.rm = TRUE)
  (x - min_val) / (max_val - min_val)
}

# Apply the normalization function using calc
temp_normalized <- calc(temp_calculation, norm_calculation)

# Now, create a new raster where the UrbanExProbability is zero
# First, make a copy of UrbanExProbability
Hotspot_Stoplight_Expansion <- raster_list$UrbanExProbability

# Set values to 0 where UrbanExProbability is 0, else use normalized values
values(Hotspot_Stoplight_Expansion) <- ifelse(values(raster_list$UrbanExProbability) == 0, 
                                               0, 
                                               values(temp_normalized))

```

interpret climate risk index by excluding zero or very near zero values and then determining the percentiles from there


# PLOTS
# Define Plot Function
```{r}
# Define plot and export function
plotRasterWithFisherBreaks <- function(raster_data, plot_title) {
  # Extract Min and Max Values
  min_val <- cellStats(raster_data, stat = 'min')
  max_val <- cellStats(raster_data, stat = 'max')

  # Check if min_val and max_val are finite
  if (is.finite(min_val) && is.finite(max_val)) {
    # Determine the number of breaks
    num_breaks <- 9  # Adjust the number of breaks as needed
    
    # Calculate Fisher breaks
    values <- getValues(raster_data)
    if (!is.null(values)) { # Ensure values were successfully retrieved
      fisher_breaks <- classIntervals(values, n = num_breaks, style = "fisher")$brks
      
      # Plot preparation without opening the jpeg device
      plot_func <- function() {
        plot(raster_data, col=inferno(length(fisher_breaks)-1), breaks=fisher_breaks, main=plot_title)
      }
      
      # Display plot in R environment
      plot_func()
      
      # Save plot to JPEG
      jpeg(paste0(plot_title, ".jpg"), width = 3840, height = 2160, res = 300)
      plot_func()  # Re-run plotting function to save to file
      dev.off()  # Close the jpeg device
    } else {
      cat("Failed to retrieve values from raster or raster contains only NA values.\n")
    }
  } else {
    cat("Raster contains non-finite or NA values only\n")
  }
}


```

```{r, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
setwd(plot_dir)

plotRasterWithFisherBreaks(raster_list$Flood_Probability, "Normalized Flood Hazard  Probability of Detectable Surface Water")
plotRasterWithFisherBreaks(raster_list$Heat_Probability, "Normalized Heat Hazard  Median of 5 Hottest Recorded Temperatures")
plotRasterWithFisherBreaks(Climate_Hazards, "Normalized Climate Hazards  Normalized Flood Hazard x Normalized Heat Hazard")
plotRasterWithFisherBreaks(raster_list$population, "Normalized Population  CIESIN High Resolution Settlement Layer")
plotRasterWithFisherBreaks(raster_list$GDP, "Normalized Gridded GDP  Global Electric Consumption revised GDP")
plotRasterWithFisherBreaks(Population_Vulnerability, "Normalized Population Vulnerability  Population x (1- GDP)")
plotRasterWithFisherBreaks(Climate_Impact_Risk, "Climate Impact Risk Index  Climate Hazards x Population Vulnerability")
plotRasterWithFisherBreaks(Climate_Impact_Risk_v2, "Climate Impact Risk Index  Climate Hazards x Population")
```

```{r, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
setwd(plot_dir)

plotRasterWithFisherBreaks(raster_list$BII, "Biodiversity Intactness Index")
plotRasterWithFisherBreaks(raster_list$LCC_Probability, "Land Cover Change Probability")
# plotRasterWithFisherBreaks(raster_list$OtherProbability, "Land Cover Change Probability")
plotRasterWithFisherBreaks(Bio_Loss_Risk, "Biodiversity Loss Risk Index  Biodiversity Intactness Index x Land Cover Change Probability")
plotRasterWithFisherBreaks(UrbanProbability, "Urban Probability Index")
plotRasterWithFisherBreaks(raster_list$UrbanExProbability, "Urban Expansion Probability Index")
```

## Hotspot Stoplight Raster
```{r, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
# Define plot and export function
STOPLIGHTplotRasterWithFisherBreaks <- function(raster_data, plot_title) {
  # Define the green-yellow-red color ramp
color_ramp <- colorRampPalette(c("#a80000", "orange", 'yellow', "#328232"))

# Extract Min and Max Values
  min_val <- cellStats(raster_data, stat = 'min')
  max_val <- cellStats(raster_data, stat = 'max')

  # Check if min_val and max_val are finite
  if (is.finite(min_val) && is.finite(max_val)) {
    # Determine the number of breaks
    num_breaks <- 255  # Adjust the number of breaks as needed
    
    # Calculate Fisher breaks
    values <- getValues(raster_data)
    if (!is.null(values)) { # Ensure values were successfully retrieved
      fisher_breaks <- classIntervals(values, n = num_breaks, style = "fisher")$brks
      
      # Plot preparation without opening the jpeg device
      plot_func <- function() {
        plot(raster_data, col=color_ramp(length(fisher_breaks)-1), breaks=fisher_breaks, main=plot_title)
      }
      
      # Display plot in R environment
      plot_func()
      
      # Save plot to JPEG
      jpeg(paste0(plot_title, ".jpg"), width = 3840, height = 2160, res = 300)
      plot_func()  # Re-run plotting function to save to file
      dev.off()  # Close the jpeg device
    } else {
      cat("Failed to retrieve values from raster or raster contains only NA values.\n")
    }
  } else {
    cat("Raster contains non-finite or NA values only\n")
  }
}

setwd(plot_dir)

STOPLIGHTplotRasterWithFisherBreaks(Hotspot_Stoplight, "Integrated Development Risk Index")
STOPLIGHTplotRasterWithFisherBreaks(Hotspot_Stoplight_Expansion, "Integrated Development Risk Index - Urban Expansion")

```
```{r}
# Function to calculate cutoff rasters
calculate_cutoff_rasters <- function(rasters, quantiles) {
  results <- list() # Initialize an empty list to store results
  
  # Loop over each raster in the list
  for (raster_name in names(rasters)) {
    raster_data <- rasters[[raster_name]]
    
    # Loop over each quantile
    for (quantile_value in quantiles) {
      cutoff <- quantile(raster_data, probs = quantile_value, na.rm = TRUE)
      
      # Calculate the raster with cutoff
      cutoff_raster <- calc(raster_data, function(x) {
        x[x <= cutoff] <- NA
        return(x)
      })
      
      # Naming the raster in the format: rastername_quantile
      result_name <- paste(raster_name, quantile_value * 100, sep = "_")
      results[[result_name]] <- cutoff_raster
    }
  }
  
  return(results)
}

#calculate threshold versions for the following rasters
raster_list_for_cutoffs <- list(
  Bio_Loss_Risk = Bio_Loss_Risk,
  Climate_Impact_Risk = Climate_Impact_Risk,
  Hotspot_Stoplight = Hotspot_Stoplight,
  Hotspot_Stoplight_Expansion = Hotspot_Stoplight_Expansion
)
# set the thresholds to calculate for each
quantile_list <- c(0.75, 0.90, 0.97, 0.95)

# Call the function
cutoff_rasters <- calculate_cutoff_rasters(raster_list_for_cutoffs, quantile_list)

# Access individual rasters like this:
# cutoff_rasters[["Bio_Loss_Risk_75"]]
# cutoff_rasters[["Hotspot_Stoplight_95"]]

plot(cutoff_rasters[["Bio_Loss_Risk_75"]])
plot(cutoff_rasters[["Climate_Impact_Risk_75"]])
plot(cutoff_rasters[["Hotspot_Stoplight_97"]])
plot(cutoff_rasters[["Hotspot_Stoplight_95"]])
```


## Data Distributions
```{r}
# Extract the values from the raster (excluding NA values)
Climate_Impact_RiskVals <- getValues(Climate_Impact_Risk)
Bio_Loss_RiskVals <- getValues(Bio_Loss_Risk)
UrbanProbabilityVals <- getValues(UrbanProbability)
HotspotStoplightVals <- getValues(Hotspot_Stoplight)

# Plot the histograms
hist(Climate_Impact_RiskVals, main = "Histogram of Climate Impact Risk Values", xlab = "Risk", ylab = "Frequency", col = "blue")
hist(Bio_Loss_RiskVals, main = "Histogram of Biodiversity Loss Risk Values", xlab = "Risk", ylab = "Frequency", col = "blue")
hist(UrbanProbabilityVals, main = "Histogram of Urban Probability Values", xlab = "Probability", ylab = "Frequency", col = "blue")
hist(HotspotStoplightVals, main = "Histogram of Hotspot Stoplight Values", xlab = "Suitability", ylab = "Frequency", col = "blue")

```

#Summarize Datasets within GAM Municipal Boundaries (WIP)
```{r}
# # Check and Align CRS
# # Check CRS of both layers
# rasterLayer <- Climate_Impact_Risk_Q4
# 
# # # Identify invalid polygons
# # invalid <- !st_is_valid(muni)
# # 
# # # Report how many are invalid
# # sum(invalid)
# # 
# # # Attempt to repair invalid polygons
# # muni[invalid, ] <- st_make_valid(muni[invalid, ])
# 
# # Reproject polygons to match raster CRS if necessary
# muni <- st_transform(muni, crs = crs(rasterLayer))
# 
# # Calculate Sum of Raster Values within Each Polygon and Normalize by Area
# # Convert sf polygons to SpatialPolygonsDataFrame for compatibility with raster package
# muni_sp <- as(muni, "Spatial")
# 
# # Create an empty vector to store the normalized sum values
# normalized_sums <- vector("numeric", length = nrow(muni))
# summary(normalized_sums)
# 
# # Loop through each polygon
# for (i in 1:length(normalized_sums)) {
#   # Mask raster with the current polygon
#   masked_raster <- mask(rasterLayer, muni_sp[i, ])
#   
#   # Calculate sum of the masked raster values
#   sum_values <- sum(values(masked_raster), na.rm = TRUE)
#   
#   # Calculate the area of the polygon (in the same units as the raster)
#   area_polygon <- st_area(muni[i, ])
#   
#   # Normalize sum by area and assign to the vector
#   normalized_sums[i] <- sum_values / as.numeric(area_polygon)
# }
# 
# # Add the normalized sums to the polygons data frame
# muni$normalized_sum <- normalized_sums
# 
# 
# # For the polygons
# print(st_bbox(muni))
# 
# # For the raster
# print(extent(rasterLayer))
```


## Export Datasets for Further Analysis and Visualization

### GeoTIFFs
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
# HOTSPOT STOPLIGHT
# Export the raster as a TIFF
setwd(outputs)

# Save the final raster
writeRaster(Hotspot_Stoplight, "Hotspot_Stoplight_2024.tif", overwrite=TRUE)
writeRaster(Hotspot_Stoplight_Expansion, "Hotspot_Stoplight_Expansion_2024.tif", overwrite=TRUE)

# INTERMEDIATE INDICES
# Save the final rasters
writeRaster(Bio_Loss_Risk, "BioRiskNorm.tif", overwrite=TRUE)
writeRaster(Climate_Impact_Risk, "AnthroRiskNorm.tif", overwrite=TRUE)
writeRaster(UrbanProbability, "UrbanProbability.tif", overwrite=TRUE)
writeRaster(raster_list$LCC_Probability, "LCC_Probability.tif", overwrite=TRUE)
writeRaster(raster_list$UrbanExProbability, "UrbanExProbability.tif", overwrite=TRUE)


# FORMATTED INPUT DATASETS
writeRaster(raster_list$BII, "BII.tif", overwrite=TRUE)
writeRaster(raster_list$population, "population.tif", overwrite=TRUE)
writeRaster(raster_list$OtherProbability, "OtherProbability.tif", overwrite=TRUE)
writeRaster(raster_list$GDP, "GDP.tif", overwrite=TRUE)
writeRaster(Population_Vulnerability, "PopVulnerability.tif", overwrite=TRUE)
writeRaster(Climate_Hazards, "Climate_Hazards.tif", overwrite=TRUE)
writeRaster(raster_list$Flood_Probability, "Flood_Hazards.tif", overwrite=TRUE)
writeRaster(raster_list$Heat_Probability, "Heat_Hazards.tif", overwrite=TRUE)

# THRESHOLD RASTERS
# Function to export threshold rasters
export_rasters <- function(rasters) {
  for (raster_name in names(rasters)) {
    # Define the file path
    file_path <- file.path(paste(raster_name, "tif", sep = "."))
    
    # Export the raster as a TIFF file
    writeRaster(rasters[[raster_name]], file_path, format = "GTiff", overwrite = TRUE)
  }
}

# Export the rasters
export_rasters(cutoff_rasters)

```

### Cloud-Optimized GeoTIFFs
```{r}
# # Specify the subdirectory for Cloud Optimized GeoTIFFs
# cog_dir <- "COG"
# 
# # Combine the paths
# full_cog_path <- file.path(outputs, cog_dir)
# 
# # Check if the directory exists, if not create it
# if (!dir.exists(full_cog_path)) {
#   dir.create(full_cog_path, recursive = TRUE)
# }
# 
# # Set the working directory to the COG directory
# setwd(full_cog_path)
# 
# # Function to write a raster as a cloud-optimized GeoTIFF and create overviews
# write_cog <- function(raster_data, file_name) {
#   cog_path <- file.path(getwd(), file_name)
#   
#   # Ensure the raster data is a stars object
#   if (!inherits(raster_data, "stars")) {
#     raster_data <- st_as_stars(raster_data)
#   }
#   
#   # Write the initial GeoTIFF
#   write_stars(raster_data, cog_path, driver = "GTiff", options = c("TILED=YES", "COMPRESS=LZW", "PREDICTOR=2", "BIGTIFF=IF_SAFER"))
# 
#   # Add overviews using a system call to GDAL
#   # This assumes that 'gdaladdo' is accessible from the command line
#   command <- sprintf("gdaladdo -r average %s 2 4 8 16", shQuote(cog_path))
#   system(command)
# }
# 
# # Ensure that 'Hotspot_Stoplight' is either a stars object or a raster object that can be converted
# write_cog(Hotspot_Stoplight, "Hotspot_Stoplight_2024.tif")
# write_cog(Hotspot_Stoplight_Expansion, "Hotspot_Stoplight_Expansion_2024.tif")
# 
# # Export other datasets as COG
# write_cog(Bio_Loss_Risk, "BioRiskNorm.tif")
# write_cog(Climate_Impact_Risk, "AnthroRiskNorm.tif")
# write_cog(UrbanProbability, "UrbanProbability.tif")
# write_cog(raster_list$LCC_Probability, "LCC_Probability.tif")
# 
# # Export more datasets as needed using the same function...
# 
# # Function to export a list of raster datasets as COGs
# export_cog_rasters <- function(rasters) {
#   for (raster_name in names(rasters)) {
#     # Ensure each raster is in an appropriate format for conversion to stars object
#     raster_data <- rasters[[raster_name]]
#     cog_filename <- paste0(raster_name, ".tif")
#     
#     # Export the raster as a COG file
#     write_cog(raster_data, cog_filename)
#   }
# }
# 
# # Export threshold rasters as COG
# export_cog_rasters(cutoff_rasters)


```

