---
title: "Raster Pre-Processing"
author: "Oliver Atwood"
output: html_document
date: "`r Sys.Date()`"
---
# Overview
<br>
This r markdown script pre-processes raster data used in the Hotspot Stoplight Project for visualization and cluster analysis. The datasets used in this project are as follows:

* LAND COVER CHANGE PROBABILITY - [Clark Labs](https://env1.arcgis.com/arcgis/rest/services/Land_Cover_Vulnerability_2050/ImageServer0)
* BIODIVERSITY INTACTNESS INDEX - [Impact Observatory](https://gee-community-catalog.org/projects/bii/)
* CLIMATE RISK INDEX - Hotspot Stoplight Team: [Nissim Lebovits](http://nlebovits.github.io/)
* GRIDDED 2020 POPULATION - [Global Human Settlements Layer](https://ghsl.jrc.ec.europa.eu/download.php?ds=bu)
* URBAN LAND COVER PROBABILITY - Hotspot Stoplight Team: [Oliver Atwood](https://www.linkedin.com/in/oliveratwood), using classified data from [Tristan Grupp](https://scholar.google.com/citations?user=yX3Ed8EAAAAJ&hl=en)
<br>
<br>

* Each of these raster datasets are loaded into R and their Coordinate Systems, Resolution, and Extent are standardized. NA values are then removed and each dataset is standardized.
<br>
* Four of these standardized datasets are then combined to produce two 'Interaction Rasters', using the following formulas:
<br>
+ Bio_Loss_Risk = LCC_Probability * BII
<br>
*Bio_Loss_Risk emphasizes areas of high land cover change probability and high biodiversity intactness.*
<br>
+ Climate_Impact_Risk = Climate_Hazards * population
<br>
*Climate_Impact_Risk emphasizes areas of high climate hazard probability and high human population.*
<br>
+ Urban Land Cover Probability is a dataset generated through a random forest model trained on a range of physiographic factors to predict the likelihood of a given cell of a land cover raster to be urban in 2033.
<br>

* Bio_Loss_Risk and Climate_Impact_Risk were both normalized and Bio_Loss_Risk, Climate_Impact_Risk, and Urban Land Cover Probability were exported for clustering analysis.
<br>
* These three datasets were then added together and the resultant raster dataset was normalized to produce a 'Stoplight for Urban Development'.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE)
rm(list=ls())
```

```{r, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide', echo=FALSE}
library(sf)
library(raster)
library(dplyr)
library(readr)
library(terra)
library(spdep)
library(rasterVis)
library(lattice)
library(gridExtra)
library(viridis)
library(classInt)

# Capture the start time
# start_time <- Sys.time()
```

## Setting Parameters
```{r, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide', echo=FALSE}
coordinate_system <- 4326
# Target resolution in m
resolution <- 30
# resolution <- 100
# Conversion to degrees
resolution <- (resolution/30)/3600

```

## Loading Datasets
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE, results = 'hide'}

# inputs <- '/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/inputs/cropped/crop4_100m'
# inputs <- '/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/inputs/cropped/crop6_30m'
inputs <- '/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/inputs/cropped/crop5c_60m'

# outputs <- "/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop4_100m"
# outputs <- "/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop6_30m"
outputs <- "/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop5c_60m"


# Base directory
inputs <- '/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/inputs/'

# # List of file names
# filenames <- c('lc_2023.tif', 'lc_2050.tif', 'Built.tif', 'Other.tif')
# var_names <- c('lc_2023', 'lc_2050', 'UrbanProbability', 'OtherProbability')
# 
# # Loop through each file name and assign its full path to a dynamically named variable
# for (i in 1:length(filenames)) {
#   # Generate the full path for the current file
#   full_path <- file.path(inputs, filenames[i])
# 
#   # Dynamically create a variable with the name from var_names and assign the full path to it
#   assign(var_names[i], full_path)
# }

# lc_2023 <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop4_100m/lc_2023.tif")
# lc_2050 <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop4_100m/lc_2050.tif")
# UrbanProbability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop4_100m/Built.tif")
# OtherProbability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop4_100m/Other.tif")

# lc_2023 <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop6_30m/lc_2023.tif")
# lc_2050 <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop6_30m/lc_2050.tif")
# UrbanProbability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop6_30m/Built.tif")
# OtherProbability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop6_30m/Other.tif")

lc_2023 <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop5c_60m/lc_2023.tif")
lc_2050 <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop5c_60m/lc_2050.tif")
UrbanProbability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop5c_60m/Built.tif")
OtherProbability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/LandCoverChange/data/outputs/cropped/crop5c_60m/Other.tif")

BII <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Biodiversity/data/outputs/BiodiversityIntactness_2020.tif")
Flood_Probability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/outputs/data-costa_rica-outputs-flood_prob.tif")
Heat_Probability <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/outputs/data-costa_rica-outputs-predicted_hot_days_costa_rica.tif")
population <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/vulnerability/data/inputs/High_Resolution_Settlement_Layer_crop2.tif")
GDP <- raster("/Users/oliveratwood/Documents/GitHub/HotspotStoplight/Climate/vulnerability/data/inputs/GDP_2018_USD_Millions_Crop2.tif")


# Read the Boundary file
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5a.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5b.geojson")
cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop5c.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop6.geojson")
# cropbox <- st_read("https://raw.githubusercontent.com/HotspotStoplight/HotspotStoplight/main/CropBoxes/CR_Crop4.geojson")

# # PRESENT LAND COVER
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/lc_2023.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# lc_2023 <- raster(temp)
# 
# # PREDICTED LAND COVER
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/lc_2050.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# lc_2050 <- raster(temp)
# 
# # LAND COVER CHANGE PROBABILITY
# # CLARK LABS
# # url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/Land_Cover_Change_Probs_Clark.tif"
# 
# # URBAN
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/Built.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# UrbanProbability <- raster(temp)
# 
# # OTHER
# url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/outputs/cropped/crop4_100m/Other.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# OtherProbability <- raster(temp)
# 
# # LCC_Probability <- OtherProbability + UrbanProbability
# # LCC_Probability <- OtherProbability
# 
# # BIODIVERSITY INTACTNESS INDEX
# url <- "https://github.com/HotspotStoplight/Biodiversity/raw/main/data/outputs/BiodiversityIntactness_2020.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# BII <- raster(temp)
# 
# # CLIMATE HAZARDS
# # Flood Hazard Probability
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/outputs/data-San_Jose-outputs-flood_prob.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# Flood_Probability <- raster(temp)
# 
# # Heat Hazard Probability
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/outputs/data-costa_rica-outputs-predicted_hot_days_costa_rica.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# Heat_Probability <- raster(temp)
# 
# # POPULATION
# # url <- "https://github.com/HotspotStoplight/LandCoverChange/raw/main/data/Pop_2020_Crop4.tif"
# # temp <- tempfile(fileext = ".tif")
# # download.file(url, temp, method = "auto")
# # population <- raster(temp)
# 
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/vulnerability/data/inputs/High_Resolution_Settlement_Layer_crop2.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# population <- raster(temp)
# 
# # GDP
# url <- "https://github.com/HotspotStoplight/Climate/raw/main/vulnerability/data/inputs/GDP_2018_USD_Millions_Crop2.tif"
# temp <- tempfile(fileext = ".tif")
# download.file(url, temp, method = "auto")
# GDP <- raster(temp)

# List of raster variables
raster_list = list(BII, Flood_Probability, Heat_Probability, population, UrbanProbability, OtherProbability, lc_2023, lc_2050, GDP)

```
## Define Normalization Functions
MIN-MAX
```{r}
norm <- function(raster) {
  # Calculate the minimum and maximum values of the raster
  min_val <- minValue(raster)
  max_val <- maxValue(raster)
  
  # Perform min-max normalization
  # Note: Use overlay or calc function for raster operations
  normalized_raster <- calc(raster, function(x) {
    # Apply the min-max normalization formula
    (x - min_val) / (max_val - min_val)
  })
  
  # Handle any potential division by zero or other anomalies
  normalized_raster[is.na(normalized_raster) | is.infinite(normalized_raster)] <- NA
  
  return(normalized_raster)
}
```

## Unify Raster Coordinate Systems, Resolutions, and Extents
```{r}
# Extract extent and crs from cropbox
bbox <- st_bbox(cropbox)
extent <- as(extent(bbox), "Extent")
crs <- crs(cropbox)
res <- resolution

# Calculate the number of columns and rows based on the resolution and the extent
nrows <- ceiling((extent@ymax - extent@ymin) / res)
ncols <- ceiling((extent@xmax - extent@xmin) / res)

# Create an empty raster with the specified extent, resolution, and CRS
# Here, we manually specify xmn, xmx, ymn, ymx based on the extent object
r <- raster(nrows=nrows, ncols=ncols, xmn=extent@xmin, xmx=extent@xmax, ymn=extent@ymin, ymx=extent@ymax, crs=crs)

# Define the target characteristics from the raster 'r'
target_crs <- crs(r)
target_extent <- extent(r)
target_resolution <- res(r)

# Initialize an empty list to store the standardized rasters
standardized_rasters <- list()

# Loop through each raster in raster_list to standardize them
for (i in 1:length(raster_list)) {
  # Read the raster
  current_raster <- raster_list[[i]]
  
  # Step 1: Reproject the raster to match the target CRS, using compareCRS for comparison
  if (!compareCRS(crs(current_raster), target_crs)) {
    current_raster <- projectRaster(current_raster, crs = target_crs)
  }
  
  # Step 2: Resample the raster to match the target resolution and extent using bilinear interpolation
  current_raster <- resample(current_raster, r, method = "bilinear")
  
  # Step 3: Crop the raster to ensure it matches the target extent exactly
  current_raster <- crop(current_raster, target_extent)
  
  # Add the standardized raster to the list
  standardized_rasters[[i]] <- current_raster
}

# Naming the lists after projection
raster_names <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "lc_2023", "lc_2050", "GDP")
names(standardized_rasters) <- raster_names

# Overwrite raster_list with the standardized rasters
raster_list <- standardized_rasters
```


## Remove all NA values from rasters
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
# Function to load rasters, replace NA with 0, and return the modified list
process_rasters <- function(raster_list) {
  processed_list <- list() # Initialize an empty list to store processed rasters
  
  for (i in 1:length(raster_list)) {
    # Copy the raster to avoid modifying the original list directly
    raster_data <- raster_list[[i]]
    
    # Replace NA values with 0
    raster_data[is.na(raster_data)] <- 0
    
    # Add the processed raster to the new list
    processed_list[[i]] <- raster_data
  }
  
  # Return the processed list of rasters
  return(processed_list)
}

# Apply the function to raster_list
raster_list <- process_rasters(raster_list)

# Naming the processed list as per the original names
names(raster_list) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "lc_2023", "lc_2050", "GDP")

```


```{r, fig.width=10}
lc_2023 <- raster_list$lc_2023
lc_2050 <- raster_list$lc_2050

# Initialize lc_change with the same dimensions as lc_2023, filled with zeros
lc_change <- setValues(lc_2023, getValues(lc_2023) * 0)

# Applying conditions for lc_change
values(lc_change)[values(lc_2023) == 1 & values(lc_2050) == 1] <- 0 # if 2023 is forest and 2050 is forest, no change
values(lc_change)[values(lc_2023) == 2 & (values(lc_2050) == 1 | values(lc_2050) == 2)] <- 0 # if 2023 is other and 2050 is forest or other, no change
values(lc_change)[(values(lc_2023) == 3 | values(lc_2023) == 4)] <- 0 # if 2023 is urban or water, no change
values(lc_change)[values(lc_2023) == 1 & values(lc_2050) == 2] <- 2 # if 2023 is forest and 2050 is other, change to other
values(lc_change)[(values(lc_2023) == 2 & values(lc_2050) == 3) | (values(lc_2023) == 1 & values(lc_2050) == 3)] <- 1 # if 2023 is forest or other and 2050 is urban, change to urban

# Initialize LCC_Probability with UrbanProbability dimensions and values
LCC_Probability <- (lc_change * 0)

# # Assigning penalized probabilities for non-urban, non-other pixels
# penalty_factor <- 0.1
# 
# # For cells where the change is not urban - assign half probability
# nonUrbanChangeIndices <- which(values(lc_change) != 1) # Get the indices of the cells where the change is not urban
# values(LCC_Probability)[nonUrbanChangeIndices] <- values(UrbanProbability)[nonUrbanChangeIndices] * penalty_factor # Adjust these probabilities
# 
# # For cells where the change is not other - assign half probability
# nonOtherChangeIndices <- which(values(lc_change) != 2) # Get the indices of the cells where the change is not urban
# values(LCC_Probability)[nonOtherChangeIndices] <- values(OtherProbability)[nonOtherChangeIndices] * penalty_factor # Adjust these probabilities


# # Log-Adjusted Probabilities (not predicted change class)
# # Process for non-urban changes
# nonUrbanChangeIndices <- which(values(lc_change) != 1)  # Indices where change is not urban
# urbanProbValues <- values(UrbanProbability)[nonUrbanChangeIndices]  # Urban probabilities at those indices
# urbanProbValues[urbanProbValues > 0] <- log(urbanProbValues[urbanProbValues > 0])  # Log transform non-zero values
# values(LCC_Probability)[nonUrbanChangeIndices] <- urbanProbValues  # Update LCC_Probability for non-urban indices
# 
# # Process for non-other changes
# nonOtherChangeIndices <- which(values(lc_change) != 2)  # Indices where change is not other
# otherProbValues <- values(OtherProbability)[nonOtherChangeIndices]  # Other probabilities at those indices
# otherProbValues[otherProbValues > 0] <- log(otherProbValues[otherProbValues > 0])  # Log transform non-zero values
# values(LCC_Probability)[nonOtherChangeIndices] <- otherProbValues  # Correctly update LCC_Probability for non-other indices



## Assigning probabilities by predicted land cover change class
# For changes to urban, use the precise indexing approach
urbanChangeIndices <- which(values(lc_change) == 1) # Get the indices of the cells where the change is urban
values(LCC_Probability)[urbanChangeIndices] <- values(UrbanProbability)[urbanChangeIndices] # Assign the probability of urban change to the corresponding cells

# For changes to other, apply the same indexing approach
otherChangeIndices <- which(values(lc_change) == 2) # Get the indices of the cells where the change is other
values(LCC_Probability)[otherChangeIndices] <- values(OtherProbability)[otherChangeIndices] # Assign the probability of other change to the corresponding cells


# Calculating change_binary
# change_binary <- LCC_Probability # Initialize change_binary with LCC_Probability
# change_binary[getValues(LCC_Probability) > 0] <- 1 # if LCC_Probability > 0, assign 1
# change_binary[getValues(LCC_Probability) <= 0] <- 0 # if LCC_Probability <= 0, assign 0

# Add LCC_Probability to the raster_list
raster_list$LCC_Probability <- LCC_Probability

# Optionally, if you want to keep the names consistent and in a specific order, you can rename all at once
names(raster_list) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "lc_2023", "lc_2050", "GDP", "LCC_Probability")

```
```{r}
plot(LCC_Probability, col=viridis::viridis(100))
```


## Normalize all rasters
```{r}
# Apply the normalization function to each raster layer in the list
normalized_layers <- lapply(raster_list, function(layer) {
  norm(layer)
})
raster_list <- normalized_layers
names(raster_list) <- c("BII", "Flood_Probability", "Heat_Probability", "population", "UrbanProbability", "OtherProbability", "lc_2023", "lc_2050", "GDP", "LCC_Probability")
```

## Plot Input Rasters
```{r, fig.width=10}
# Convert the Raster list to a stack
raster_stack <- stack(normalized_layers)
plot(raster_stack, col=viridis::viridis(100))
```

Plot Histograms of Input Rasters
```{r, fig.width=6}
# Assuming 'raster_stack' is your RasterStack
num_layers <- nlayers(raster_stack)

# Get the names of the layers in the raster stack
layer_names <- names(raster_stack)

# Define how many histograms to plot per figure
histograms_per_figure <- 10  # Adjust based on your preference and screen size

# Calculate the number of figures needed
num_figures <- ceiling(num_layers / histograms_per_figure)

# Loop through each figure
for (fig in 1:num_figures) {
  # Setup the plotting area for the current figure
  par(mfrow=c(5, 2), mar=c(2, 2, 2, 2))  # Adjust 'mfrow' and 'mar' as needed
  
  # Calculate the index of the first and last histogram to plot in the current figure
  first_hist <- (fig - 1) * histograms_per_figure + 1
  last_hist <- min(fig * histograms_per_figure, num_layers)
  
  # Loop through the histograms to plot in the current figure
  for (i in first_hist:last_hist) {
    # Extract the layer
    layer <- raster_stack[[i]]
    
    # Extract values from the layer, removing NAs
    values <- getValues(layer)
    
    # Plot histogram of the values
    hist(values, main=layer_names[i], xlab="Value", ylab="Frequency")
  }}


```
```{r}
# # Use sapply to apply the 'res' function to each raster in the list and capture the output
# resolutions <- sapply(raster_list, res)
# extent <- sapply(raster_list, extent)
# 
# # Print the resolutions
# print(resolutions)
# print(extent)
```

## Generate Interaction Rasters
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
#BIO RISK
# Multiply LCC_Probability and BII
# Bio_Loss_Risk <- raster_list$BII * raster_list$LCC_Probability
Bio_Loss_Risk <- raster_list$BII * raster_list$OtherProbability

#CLIMATE RISK
#Generate Climate Risk Raster
Climate_Hazards <- raster_list$Flood_Probability * raster_list$Heat_Probability

#Generate vulnerability-adjusted population
# Population_Vulnerability <- (raster_list$population / raster_list$GDP)
Population_Vulnerability <- (raster_list$population)

# Multiply Vulnerability-adjusted population and Climate_Hazards
Climate_Impact_Risk <- Population_Vulnerability * Climate_Hazards

# Normalize the interaction rasters
Bio_Loss_Risk_norm <- norm(Bio_Loss_Risk)
Climate_Impact_Risk_norm <- norm(Climate_Impact_Risk)
UrbanProbability_norm <- norm(standardized_rasters$UrbanProbability)

```

## Overlay Rasters to Produce 'Stoplight For Urban Development'
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
# Weighted Overlay Calculation
Hotspot_Stoplight <- norm(UrbanProbability_norm - Bio_Loss_Risk_norm - Climate_Impact_Risk_norm)
```

interpret climate risk index by excluding zero or very near zero values and then determining the percentiles from there


# PLOTS
# Define Plot Function
```{r}
plotRasterWithFisherBreaks <- function(raster_data, plot_title) {
  # Extract Min and Max Values
  min_val <- cellStats(raster_data, stat = 'min')
  max_val <- cellStats(raster_data, stat = 'max')

  # Check if min_val and max_val are finite
  if (is.finite(min_val) && is.finite(max_val)) {
    # Determine the number of breaks
    num_breaks <- 9  # Adjust the number of breaks as needed
    
    # Calculate Fisher breaks
    values <- getValues(raster_data)
    if (!is.null(values)) { # Ensure values were successfully retrieved
      fisher_breaks <- classIntervals(values, n = num_breaks, style = "fisher")$brks
      
      # Plot the raster using the defined color ramp and Fisher breaks
      plot(raster_data, col=inferno(length(fisher_breaks)-1), breaks=fisher_breaks, main=plot_title)
    } else {
      cat("Failed to retrieve values from raster or raster contains only NA values.\n")
    }
  } else {
    cat("Raster contains non-finite or NA values only\n")
  }
}
```

```{r, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
plotRasterWithFisherBreaks(raster_list$Flood_Probability, "Flood Hazard - Probability of Detectable Surface Water")
plotRasterWithFisherBreaks(raster_list$Heat_Probability, "Heat Hazard - Probability of Multiple Days Over 35C")
plotRasterWithFisherBreaks(Climate_Hazards, "Climate Hazards - Flood and Heat Probability")
plotRasterWithFisherBreaks(raster_list$population, "Population - CIESIN High Resolution Settlement Layer")
plotRasterWithFisherBreaks(Climate_Impact_Risk_norm, "Climate Impact Risk Index")

# plotRasterWithFisherBreaks(raster_list$GDP, "Gridded GDP - Global Electric Consumption revised GDP")
# plotRasterWithFisherBreaks(Population_Vulnerability, "Population Vulnerability - Population / GDP")

```
```{r, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}

plotRasterWithFisherBreaks(raster_list$BII, "Biodiversity Intactness Index")
# plotRasterWithFisherBreaks(raster_list$LCC_Probability, "Land Cover Change Probability")
plotRasterWithFisherBreaks(raster_list$OtherProbability, "Land Cover Change Probability")
plotRasterWithFisherBreaks(Bio_Loss_Risk_norm, "Biodiversity Loss Risk Index")
plotRasterWithFisherBreaks(UrbanProbability_norm, "Urban Probability Index")

```

## Hotspot Stoplight Raster
```{r, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
STOPLIGHTplotRasterWithFisherBreaks <- function(raster_data, plot_title) {
  # Define the green-yellow-red color ramp
color_ramp <- colorRampPalette(c("red", "yellow", "darkgreen"))

# Extract Min and Max Values
  min_val <- cellStats(raster_data, stat = 'min')
  max_val <- cellStats(raster_data, stat = 'max')

  # Check if min_val and max_val are finite
  if (is.finite(min_val) && is.finite(max_val)) {
    # Determine the number of breaks
    num_breaks <- 7  # Adjust the number of breaks as needed
    
    # Calculate Fisher breaks
    values <- getValues(raster_data)
    if (!is.null(values)) { # Ensure values were successfully retrieved
      fisher_breaks <- classIntervals(values, n = num_breaks, style = "fisher")$brks
      
      # Plot the raster using the defined color ramp and Fisher breaks
      plot(raster_data, col=color_ramp(length(fisher_breaks)-1), breaks=fisher_breaks, main=plot_title)
    } else {
      cat("Failed to retrieve values from raster or raster contains only NA values.\n")
    }
  } else {
    cat("Raster contains non-finite or NA values only\n")
  }
}


STOPLIGHTplotRasterWithFisherBreaks(Hotspot_Stoplight, "Integrated Development Risk Index")

```


## Data Distributions
```{r}
# Extract the values from the raster (excluding NA values)
Climate_Impact_RiskVals <- getValues(Climate_Impact_Risk_norm)
Bio_Loss_RiskVals <- getValues(Bio_Loss_Risk_norm)
UrbanProbabilityVals <- getValues(UrbanProbability_norm)
HotspotStoplightVals <- getValues(Hotspot_Stoplight)

# Plot the histograms
hist(Climate_Impact_RiskVals, main = "Histogram of Climate Impact Risk Values", xlab = "Risk", ylab = "Frequency", col = "blue")
hist(Bio_Loss_RiskVals, main = "Histogram of Biodiversity Loss Risk Values", xlab = "Risk", ylab = "Frequency", col = "blue")
hist(UrbanProbabilityVals, main = "Histogram of Urban Probability Values", xlab = "Probability", ylab = "Frequency", col = "blue")
hist(HotspotStoplightVals, main = "Histogram of Hotspot Stoplight Values", xlab = "Suitability", ylab = "Frequency", col = "blue")

```

## Export Datasets for Further Analysis and Visualization
```{r, message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE}
# HOTSPOT STOPLIGHT
# Export the raster as a TIFF
setwd(outputs)

# Save the final raster
writeRaster(Hotspot_Stoplight, "Hotspot_Stoplight_2024.tif", overwrite=TRUE)

# INTERMEDIATE INDICES
# Save the final rasters
writeRaster(Bio_Loss_Risk_norm, "BioRiskNorm.tif", overwrite=TRUE)
writeRaster(Climate_Impact_Risk_norm, "AnthroRiskNorm.tif", overwrite=TRUE)
writeRaster(UrbanProbability_norm, "UrbanProbability.tif", overwrite=TRUE)
writeRaster(LCC_Probability, "LCC_Probability.tif", overwrite=TRUE)

```